Malacology: A Programmable Storage System

### 2017-02-07 paper/paper.tex
paper: integrate design and implementation sections
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\documentclass[preprint]{sigplanconf-eurosys}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{pifont}
\usepackage{color}
\newcommand{\cmark}{\ding{51}}%
\usepackage[font=footnotesize,labelfont=bf]{caption}

@@

\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
\usepackage{pstricks}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
\setkeys{Gin}{width=\maxwidth,keepaspectratio}

@@

@@

To address these challenges we present Malacology. Malacology is both a
prototype for a programmable storage system and a design approach to evolve
storage systems efficiently and without jeopardizing years of code-hardening
and performance optimization efforts.  Although Malacology uses the internal
abstractions of the underlying storage system, including its subsystems,
components, and implementations, we emphasize that our system still addresses
the general challenges outlined above.

@@

\begin{table*}
\centering\small
\begin{tabular}{  l | l | l | l }
%\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
\multicolumn{4}{c}{} \\
\textbf{Abstraction}                    &
\textbf{Example in Production Systems}  &
\textbf{Example in Ceph}                &
\textbf{Provided Primitives}            \\ \hline
{\it Service Metadata}
  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,burrows_chubby_2006}
  & cluster state management~\cite{website:ceph-mon}
  & consensus \& consistency
  \\
{\it Shared Resource}
  & MPI collective IO, burst
  & POSIX metadata protocols
  & serialization \& batching
  \\
{\it Object \& Data}
  & Swift in situ storage/compute~\cite{website:zerocloud}
  & object interface classes~\cite{website:cls-lua}
  & transaction \& atomicity
  \\
Durability
  & S3/Swift interfaces (RESTful API)
  & object store library~\cite{weil_rados_2007}
  & interface to persistence
  \\
Load Balancing
  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
  & migration \& sampling
  \\
\end{tabular}
\caption{Common internal abstractions. The same ``internal abstractions" are common across large-scale
systems because they provide primitives that solve general distributed systems
problems.  Here we list examples of what these internal abstractions are used
for in ``production systems" and in Ceph.  Malacology provides these internal
abstractions as interfaces that higher level applications can use; {\it
italicized abstractions} are interfaces discussed in this paper.}
\label{table:examples}
\end{table*}

The guiding principle
is to re-use existing services and extend them so that these services can be
\emph{programmed}. We accomplish programmability of a service by exporting
bindings (or ``hooks'') for an interpreted programming language so that
programming can occur without having to restart the storage system (see also
below,~\S\ref{sec:durability}). Table~\ref{table:examples} shows the internal services from Ceph that we expose in the Malacology prototype via Lua~\cite{ierusalimschy1996lua} bindings.

@@

In the case of Ceph, a consistent view of cluster state among server daemons
and clients is critical to provide strong consistency guarantees to clients.
Ceph maintains cluster state information in per-subsystem ({\it e.g.} the
metadata server) data structures called ``maps'' that record membership and
status information.  A Paxos~\cite{lamport_parttime_1998} monitoring service is
responsible for integrating state changes into cluster maps, responding to
requests from out-of-date clients and synchronizing members of the cluster
whenever there is a change in a map so that they all observe the same system
state. As a fundamental building block of many system designs, consensus
abstractions such as Paxos are a common technique for maintaining consistent
data versions, and are a useful system to expose.

@@

An object interface is a plugin structured similarly to that of an RPC
in which a developer creates a named function within the cluster that clients
may invoke. In the case of Ceph each function is implicitly run within the
context of an object specified when the function is called. Developers of
object interfaces express behavior by creating a composition of native
interfaces or other custom object interfaces, and handle serialization of
function input and output.  A wide range of native interfaces are available to
developers such as reading and writing to a byte stream, controlling object
snapshots and clones, and accessing a sorted key-value database. These native
interfaces may be transactionally composed along with application
specific logic to create semantically rich interfaces. An example would be
an interface that atomically updates a matrix stored in the bytestream and an
index of the matrix stored in the key-value database.

@@

The distributed metadata service in Ceph provides clients with a POSIX file
system abstraction~\cite{weil:sc2004-dyn-metadata}. In general, distributed
file systems protect resources by providing hierarchical indexing and
distributed locking services.  In Ceph, the locking service implements a
capability-based system that expresses what data and metadata clients are
allowed to access as well as what state they may cache and modify locally.
While designed for the file abstraction, indexing, locking, and caching are
all common services that are useful to a broad spectrum of applications.
Distributed applications that share centralized resources (e.g. a database or
directory) face similar challenges which are often solved using
application-specific sharding.

>>>>

\documentclass[preprint]{sigplanconf-eurosys}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{pifont}
\usepackage{color}
\usepackage{comment}
\includecomment{newcomment}
\includecomment{oldcomment}
\newcommand{\cmark}{\ding{51}}%
\usepackage[font=footnotesize,labelfont=bf]{caption}

@@

\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
\usepackage[normalem]{ulem}
\newcommand{\nc}[1]{{
  \textcolor{red}{\textbf{#1}}
}}
\newcommand{\oc}[1]{{
  \textcolor{blue}{\textbf{\sout{#1}}}
}}

\usepackage{pstricks}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
\setkeys{Gin}{width=\maxwidth,keepaspectratio}

@@

\begin{newcomment}

\nc{New changes are in red}

\oc{Old prose is in blue}

%\nc{
%\begin{enumerate}
%  \item clarify what was implemented: diagram of interfaces and prose specifying the difference between Malacology and prior work (reviewers B, C, D, E)
%  \item describe what makes a programmable storage system (reviewers C, D)
%  \item organize motivation: group interfaces into categories (e.g., core functionality, features, and performance optimizations) and describe other applications we can build with Malacology (reviewers B, C)
%  \item clarify terminology: service vs. file system metadata, storage vs. file system, dirty vs. clean slate approaches, soft-state caching (reviewers B, C)
%  \item address security, access control, and how it affects composability/multi-tenancy on correctness (reviewer D)
%\end{enumerate}
%}
\end{newcomment}

@@

To address these challenges we present Malacology
\begin{oldcomment}
\oc{. Malacology is both a prototype for a programmable storage system and a}
\end{oldcomment}
\begin{newcomment}
\nc{, our prototype programmable storage system. It uses the programmable storage}
\end{newcomment}
design approach to evolve
storage systems efficiently and without jeopardizing years of code-hardening
and performance optimization efforts.  Although Malacology uses the internal
abstractions of the underlying storage system, including its subsystems,
components, and implementations, we emphasize that our system still addresses
the general challenges outlined above.

@@

\begin{table*}
\centering\small
\begin{tabular}{  l | l | l | l }
%\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
\multicolumn{4}{c}{} \\
\textbf{Interface}                      &
\textbf{Example in Production Systems}  &
\textbf{Example in Ceph}                &
\textbf{Provided Primitives}            \\ \hline
{\it Service Metadata}
  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,burrows_chubby_2006}
  & cluster state management~\cite{website:ceph-mon}
  & consensus \& consistency
  \\
{\it Shared Resource}
  & MPI collective IO, burst
  & POSIX metadata protocols
  & serialization \& batching
  \\
{\it Object \& Data}
  & Swift in situ storage/compute~\cite{website:zerocloud}
  & object interface classes~\cite{website:cls-lua}
  & transaction \& atomicity
  \\
Durability
  & S3/Swift interfaces (RESTful API)
  & object store library~\cite{weil_rados_2007}
  & interface to persistence
  \\
Load Balancing
  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
  & migration \& sampling
  \\
\end{tabular}
\caption{Common internal abstractions. The same ``internal abstractions" are common across large-scale
systems because they provide primitives that solve general distributed systems
problems.  Here we list examples of what these internal abstractions are used
for in ``production systems" and in Ceph.  Malacology provides these internal
abstractions as interfaces that higher level applications can use; {\it
italicized abstractions} are interfaces discussed in this paper.}
\label{table:examples}
\end{table*}

The guiding principle is to re-use existing services and extend them so that
these services can be \emph{programmed}. We accomplish programmability of a
service by exporting bindings (or ``hooks'') for an interpreted programming
language so that programming can occur without having to restart the storage
system (see also below,~\S\ref{sec:durability}). Table~\ref{table:examples}
shows the internal services from Ceph that we expose in the Malacology
prototype via Lua~\cite{ierusalimschy1996lua} bindings
\begin{newcomment}
\nc{and Figure~\ref{fig:implementation-overview} compares what was already
present in Ceph (gray boxes) to the Malacology interfaces we added (blue
trapezoids); Section~\ref{sec:services} will describe the higher-level
services (pink arrows) we built with these interfaces.}

\begin{figure}[tbp]
\centering
\includegraphics[width=100mm]{figures/implementation-overview.png}
\caption{\nc{Malacology is implemented on the daemons and clients that run in a
Ceph cluster. The interfaces expose internal subsystems of the Ceph modules and
are used as building blocks for higher-level services.}
\label{fig:implementation-overview}}
\end{figure}
\end{newcomment}

@@

\begin{newcomment}
  \nc{\\ \noindent\it{\textbf{Ceph Implementation: }}}
\end{newcomment}
In the case of Ceph, a consistent view of cluster state among server daemons
and clients is critical to provide strong consistency guarantees to clients.
Ceph maintains cluster state information in per-subsystem ({\it e.g.} the
metadata server) data structures called ``maps'' that record membership and
status information.  A Paxos~\cite{lamport_parttime_1998} monitoring service is
responsible for integrating state changes into cluster maps, responding to
requests from out-of-date clients and synchronizing members of the cluster
whenever there is a change in a map so that they all observe the same system
state. As a fundamental building block of many system designs, consensus
abstractions such as Paxos are a common technique for maintaining consistent
data versions, and are a useful system to expose.

@@

\begin{newcomment}
  \nc{\\ \noindent\it{\textbf{Ceph Implementation: }}}
\end{newcomment}
An object interface is a plugin structured similarly to that of an RPC
in which a developer creates a named function within the cluster that clients
may invoke. In the case of Ceph each function is implicitly run within the
context of an object specified when the function is called. Developers of
object interfaces express behavior by creating a composition of native
interfaces or other custom object interfaces, and handle serialization of
function input and output.  A wide range of native interfaces are available to
developers such as reading and writing to a byte stream, controlling object
snapshots and clones, and accessing a sorted key-value database. These native
interfaces may be transactionally composed along with application
specific logic to create semantically rich interfaces. An example would be
an interface that atomically updates a matrix stored in the bytestream and an
index of the matrix stored in the key-value database.

@@

\begin{oldcomment}
\oc{The distributed metadata service in Ceph provides clients with a POSIX file
system abstraction~\cite{weil:sc2004-dyn-metadata}.}
\end{oldcomment}
\begin{newcomment}
\nc{File systems provide clients with the familiar POSIX file abstraction.
While this guarantees strong consistency it comes at the cost of scalability
and lower performance.}
\end{newcomment}
In general, distributed file systems protect resources by providing
hierarchical indexing and distributed locking services.

\begin{newcomment}
\nc{\\ \noindent\it{\textbf{Ceph Implementation: }}}
\end{newcomment}
In Ceph, the locking service implements a
capability-based system that expresses what data and metadata clients are
allowed to access as well as what state they may cache and modify locally.
While designed for the file abstraction, indexing, locking, and caching are
all common services that are useful to a broad spectrum of applications.
Distributed applications that share centralized resources (e.g. a database or
directory) face similar challenges which are often solved using
application-specific sharding.
